More than Likes: A bla bla of bla bla with bla bla

### 2023-06-15 paper.Rmd
speciale edits

---
title: 'More than Likes: A bla bla of bla bla with bla bla'
subtitle: |
  | 
  | University of Copenhagen
  | `r source("scripts/utils.R")`
  | Student IDs:
  | XXXxxx
  | XXXxxx
  | Master's Thesis
  | Number of characters: xxxxxxxxx
  | `r # maybe for word count: RmdWords("paper.Rmd")$num_words`
output:
  pdf_document:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
  html_document: 
    number_sections: yes
    toc: true
    toc_depth: 2
header-includes:
- \usepackage{amsmath} # format math stuff
- \usepackage{booktabs} # makes table nicer
- \usepackage{lastpage} # be able to count total pages
- \usepackage{fancyhdr} # header/footer
- \pagestyle{fancy} # header/footer
- \fancyhead{} # remove header
- \renewcommand{\headrulewidth}{0pt} # remove header
- \fancyfoot[CO,CE]{Page \thepage\ of \pageref{LastPage}} # add Page X of Y
- \usepackage{floatrow} # position tables and figures correctly
- \floatsetup[figure]{capposition=top} # position tables and figures correctly
- \floatsetup[table]{capposition=top} # position tables and figures correctly
- \usepackage{setspace} # controls line spacing
bibliography: references.bib # add bibliography
csl: chicago-author-date # choose chicago citation style
link-citations: yes # add links in citations
nocite: '*' # add all used citations
---

```{r setup, include=FALSE}
source("scripts/config.R")
```

\pagebreak
<!-- TC:ignore -->
```{=latex}
\setcounter{tocdepth}{2}
\tableofcontents
```
<!-- TC:endignore -->

@@

# Introduction
<!-- Word count: approx. 400  -->
<!-- Word count now: 426  -->

\onehalfspacing <!-- OBS: We use double spacing here. Could also be changed to \onehalfspacing -->

\pagebreak
# Data and Method

\pagebreak 
# Results and discussion

<!-- TC:ignore -->
```{r, echo=FALSE, results='asis'}
source("scripts/regression.R")
```
<!-- TC:endignore -->

\pagebreak
# Conclusion

@@

<!-- TC:ignore -->

>>>>

---
indent: true
output:
  pdf_document:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
header-includes:
- \usepackage{amsmath} # format math stuff
- \usepackage{booktabs} # makes table nicer
- \usepackage{multirow} # make merged cells possible
- \usepackage{dcolumn} # makes table nicer
- \usepackage{lastpage} # be able to count total pages
- \usepackage{fancyhdr} # header/footer
- \usepackage{graphicx} #
- \usepackage{wrapfig}
- \setlength\parindent{24pt}
- \usepackage[explicit]{titlesec}
- \titlespacing\subsubsection{0pt}{12pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
- \pagestyle{fancy} # header/footer
- \fancyhead{} # remove header
- \setcounter{page}{0} #  frontpage is page 0
- \fancyfoot[CO,CE]{Page \thepage\ of \pageref{LastPage}}
- \renewcommand{\headrulewidth}{0pt} # remove header
- \usepackage{floatrow} # position tables and figures correctly
- \floatsetup[figure]{capposition=top} # position tables and figures correctly
- \floatsetup[table]{capposition=top} # position tables and figures correctly
- \usepackage{setspace} # controls line spacing
bibliography: references.bib # add bibliography
csl: chicago-author-date # choose chicago citation style
#link-citations: yes # add links in citations
nocite: '@*' # add all used citations
---

```{r setup, include=FALSE}
source("scripts/initialize.R")
source("scripts/regression.R")
```

\thispagestyle{empty}

\vspace{2cm}

\includegraphics[]{images/ku_logo_no_txt.png}

\vspace{2cm}

\Large
**University of Copenhagen**

**Department of Sociology**

<!-- ```{r, echo=FALSE, out.width="100%", fig.align='right'} -->
<!-- include_graphics("images/ku_logo_no_txt.png") -->
<!-- ``` -->

\begin{centering}

\vspace{3cm}

\Large

\doublespacing
\textbf{MORE THAN LIKES: HOW NEWS CONTENT INFLUENCES AUDIENCE ENGAGEMENT ON SOCIAL MEDIA}

\vspace{1 cm}

\normalsize
\singlespacing
A Master Thesis by

\vspace{0.5 cm}

\Large

\textbf{Tobias Priesholm Gårdhus}

\normalsize
\vspace{0.5 cm}

\&

\vspace{0.5 cm}
\Large
\textbf{Mads Lang Sørensen}

\normalsize
\vspace{1.5 cm}

Supervisor: Christian Borch

Number of characters: 217,328

Date: 31\textsuperscript{th} of May 2023

\vspace{1.5 cm}

\end{centering}
\normalsize

\newpage

\onehalfspacing <!-- OBS -->

\vspace*{2cm}

```{r, echo=FALSE, fig.align='center'}
include_graphics("images/midjourney_gen.png")
```
\begin{center}
An image generated by Midjourney's image generation model using the prompt:

\noindent
\emph{“A renaissance painting of people arguing over the content of a tabloid newspaper”.}
\end{center}

@@

## Abstract {-}
This study investigates how the content of news articles influences audience engagement on social media. Addressing research gaps in previous studies, which often rely on limited or biased samples and narrow definitions of audience engagement, this study utilizes machine learning and natural language processing techniques to analyze a comprehensive dataset of news content shared on Facebook over a span of nearly two years. By collaborating with Ekstra Bladet, the largest Danish tabloid media, we are able to examine multiple different dimensions of audience engagement encompassing information selection, emotional responses, sharing, and conversations. Through an empirical analysis of the influence of 11 news content characteristics on eight audience engagement outcomes, the analysis can be summarized in three main findings. First of all, news content characterized by unexpectedness and soft news increases individualistic behaviors, such as clicking on articles, while impact and hard news increase social behaviors, such as sharing and having conversations. Secondly, the emotional framing of news content elicits similar emotional responses among the audience, with positivity and negativity prompting more recognizing and offensive language, respectively, in the comment section. Thirdly, news content related to timeliness and proximity shows limited effects on audience engagement. These findings highlight the multidimensional nature of audience engagement, underscoring that there is no universal influence of news content on audience engagement on social media.

## Declaration of Responsibilities {-}
Tobias Priesholm Gårdhus: 1, 2.1, 2.3, 2.5, 2.6, 3.2, 3.4, 3.6, 4.1, 4.3, 4.5, 5.1, 5.3, 6

\noindent
Mads Lang Sørensen: 1, 2.2, 2.4, 2.6, 3.1, 3.3, 3.5, 3.6, 4,2, 4.4, 4.5, 5.2, 5.4, 6

## Declaration of Third-party Collaboration {-}
As part of the study, the authors collaborated with a third-party, Ekstra Bladet, for data availability. However, no conflict of interest has arisen throughout the process.

@@

\singlespacing

```{=latex}
\setcounter{tocdepth}{3}
\tableofcontents
```

\onehalfspacing <!-- OBS -->

# Introduction

## Background
News media have always played a central role in a well-functioning democracy. The democratic role of media is to serve as an independent and objective reporter of the truth (Herman & Chomsky 1988:55). A common metaphor is that the media represents the watchdog of democracy by exposing the elite and those in power (Kovach & Rosenstiel 2014:160ff). Furthermore, news media has an important obligation to inform and facilitate public discussion about important societal topics (Aalberg & Curran 2012:3; Habermas 1999). However, in recent years, the role of the media has been questioned due to new digital technologies and possibilities. Perhaps the most significant consequence is the changing relationship between media and their audiences (Richardson & Stanyer 2011; Livingstone 2016:2).

Increasingly, many people turn to social media as their preferred platform for reading and engaging with news content. For the first time in 2022, Danish audiences spent more time on social media than on traditional print, radio, and broadcast media (DR Medieforskning 2022:14). Social media allow users to engage with news content in far more ways than when it was distributed through newspapers, radio, or TV (Jenkins et al. 2013:6ff). Furthermore, media organizations, editors, and journalists no longer hold the monopoly of media production and dissemination; anyone can participate in debates and act as a reporter or commentator (Deuze 2005:451). In many ways, the audiences play a more active role in the relationship with the media. Unlike the days of opinion letters and face-to-face conversations, social media and website analytics allow audience feedback to be tracked and analyzed directly within the newsroom (Richardson & Stanyer 2011; Christin 2018; Ferrer-Conill & Tandoc 2018:436).

The audience engagement can also be seen as a part of a digital public sphere, where individuals discuss news content and exchange arguments. However, this deliberative potential may not always be realized, and some argue that the audience engagement pollutes the public debate more than they benefit it (Papacharissi 2002). As a consequence, media organizations have been shutting down the comment sections on their own websites, and instead used social media platforms such as Twitter and Facebook for social interactions (Reimer et al. 2021:4). Recently, multiple Danish media have taken a step further and decided to remove the possibility to comment on shared articles on their Facebook pages, due to the number of uncivil comments. However, this decision is not without consequences. In addition to hindering the possibility of deliberation, these media sites’ posts also reach a smaller audience – reducing the impact of the news content (Albrecht 2022). Rather than completely shutting down conversations around news content, it might be more sustainable to examine what drives different kinds of audience engagement. Following these considerations, there are democratic, societal, and academic reasons to better understand audience engagement with news content on social media.

## News Content and Audience Engagement
It is far from all kinds of news content that audiences engage with on social media. While a minority of posts receive thousands or even millions of clicks, likes, shares, comments, or other types of engagements, most posts only receive a few interactions. Understanding what drives audiences to engage or abstain from engaging with news content on social media is important. On the one hand, audience engagement is an important part of opinion-making, which is a central part of a democratic society (Steensen, Ferrer-Conill & Peters 2020:1665). On the other hand, audience engagement may also affect the production of media content by determining what content is popular and thus also what content is most profitable for the media (Martin 2019). The concept of audience engagement has been defined in many different ways, but generally it refers to an active or intentional orientation towards the news content (Broersma 2019:1). In this study, we argue that audience engagement should be understood as a multifaceted phenomenon encompassing different dimensions related to information selection, emotional reactions, sharing of content, and not least having conversations about content.

Research has identified several factors that influence audience engagement on social media. In a review by Kümpel et al. (2015) three types of factors are identified: users, content, and social networks. This distinction is reminiscent of the classical work on media and audiences by Katz & Lazarsfeld (1955), where they identify five intervening variables between media and audiences: exposure, medium, content, predispositions, and interpersonal relations. In this study, we will focus only on how one of these variables influence audience engagement: the news content itself.

To determine the characteristics of news content that might influence audience engagement, many studies have used the idea of news values. Taxonomies of different news values have traditionally been used as a journalistic guideline of a story’s newsworthiness, but it has also been adapted to explain how audiences evaluate the relevance of content (Harcup & O’Neill 2001:7; Eilders 2006). From the original formulation by Galtung & Ruge (1965) to contemporary reformulations of news value theory, the categories of news values have varied greatly, but they often include news content characteristics such as timeliness, unexpectedness, proximity, negativity, and positivity.

## Empirical Gap
The question of how news content influences audience engagement on social media has been approached in various fields within both social sciences, humanities, and natural sciences (Kümpel et al. 2015:8). Although a lot of research has been conducted within this field, we argue that specifically two empirical gaps remain.

First and foremost, most studies have relied on smaller samples of the shared news content. Articles, posts, and comments have mainly been sampled based on different analytical heuristics (e.g., selection for qualitative analysis) or technical limitations (e.g., the restrictiveness of social media platform’s data access), while few studies have computationally analyzed a full population of posts and comments over an extended period. Furthermore, some studies have only studied the news content that is most engaged with, but not what is less engaged with (Kümpel et al. 2015:8-9). This limitation often stems from data constraints, where empirical analyses are restricted to the top news articles and their engagement. This approach is particularly problematic, since it only examines what is successful and not what is not successful. To better determine the factors driving audience engagement, the complete population of news articles should be included in the analysis.

A second limitation of previous research is the predominantly simple and narrow conceptualization and measurement of audience engagement. Often, audience engagement is aggregated into basic popularity cues, such as the number of likes or shares (Steensen et al. 2020:1663). However, audience engagement is a multifaceted phenomenon that can be both behavioral and felt (Steensen et al. 2020:1665). While the felt engagement is difficult to capture with quantitative metrics, digital traces enable us to construct different measurements for behavior. Even if we only consider behavioral engagement, there might be a significant difference between a user simply clicking the link, liking a post, sharing it to their own feed or even discussing the content in a comment thread. While quantitative popularity cues are often the focus for media organizations due to their connection to ad revenue, audience engagement may also contain undesired behaviors such as incivilities in the comment sections (Martin 2019:21; Su et al. 2018; Chen et al. 2020). To our knowledge, no previous empirical studies have examined how different news content might drive different dimensions of audience engagement.

## Research Question
This study seeks to close the empirical gaps described in the previous section. Overall, we seek to answer the following research question:

>*How does the content of news articles influence audience engagement on social media?*

## Analytical Contribution
This study aims to bridge the empirical gaps in previous research on the relation between audience engagement and news content by constructing a unique and comprehensive dataset. The creation of our dataset and analysis is made possible by collaborating with Ekstra Bladet - the largest tabloid media in Denmark. Approximately one million people visit their website every day. Ekstra Bladet also publish a printed newspaper and share their articles on various social media such as their Facebook page with over 400,000 followers. While tabloid in style, they cover all areas of society and journalistic practice, ranging from celebrity gossip to award-winning investigative journalism (Journalistforbundet 2023).

By collaborating with Ekstra Bladet, we can construct a more complete dataset that overcomes the empirical limitations of previous research. First of all, we are able to include all articles shared on Ekstra Bladet’s Facebook page for almost two years in our analysis. This allows us to analyze a complete population of articles for a longer period – rather than relying on a smaller sample that could contain biases or lack representativeness. To be able to construct measures of audience engagement and news content characteristics for this extensive dataset, we apply computational methods, including machine learning and natural language processing techniques, rather than the manual hand-labelling that most previous studied have relied upon. Secondly, we examine social media engagement by a multitude of different manifestations instead of only relying on a simple and narrow measurement. Our access to Ekstra Bladet’s Facebook page allows us to retrieve detailed post and comment data from Facebook, and therefore collect a multitude of features not otherwise available. Instead of only retrieving common popularity cues such as the number of likes, we are also able to collect text content and structure of the comment section, and metrics that are only visible to the Facebook page-owner i.e., number of clicks on the shared link. In this way, our study extends and improves the previous empirical foundation by both including a more complete population of news content and a more comprehensive understanding of audience engagement on social media.

## Data and Methods
Our dataset is constructed by combining two different data sources. The first dataset comes from Ekstra Bladet’s database and consists of article content and aggregated user behavior on their website. The second dataset is collected from Facebook and consists of all posts made by Ekstra Bladet’s Facebook page, along with information about the engagement for the posts and their comment sections. Our unit of observation is therefore the shared posts for which we construct the features of news content characteristics and audience engagement on Facebook.

The dataset spans over almost two years from the 28^th^ of February 2021 to the 1^st^ of January 2023 and contains 18,350 Facebook posts. In aggregate metrics, our measures of audience engagement encompass approximately 5.5 million comments, 7 million likes, and 230 million clicks on links to the shared articles. As audience engagement is defined as a multifaceted phenomenon, we operationalize its various dimensions on social media into eight different outcomes related to information selection, emotional responses, sharing, and conversations about news content. Similarly, we define 11 news content characteristics mainly derived from news value theory. Utilizing multiple linear regression, we analyze the influence of these 11 news content characteristics on each of the eight measurements of audience engagement on social media.

## Reading Guide
The thesis is structured as follows. Section 2 (Theoretical Background) delves into the existing body of knowledge surrounding audience engagement and news content. We review relevant previous studies, theories, and frameworks that inform our research, and highlight the two research gaps that our study aims to address. Section 3 (Data and Methods) describes the research design, data collection, operationalization of audience engagement and news content, and the analytical methods employed. The last part of Section 3 also outlines important background information about our case of study, Ekstra Bladet. Section 4 (Analysis) presents the results of our empirical analysis structured in four dimensions of audience engagement on social media: information selection, emotional responses, sharing, and conversations. Concluding on Section 4, we present three main findings across the outcomes of audience engagement on social media. In Section 5 (Discussion), we discuss our empirical research in the light of the societal and theoretical implications, the reliability and validity, as well as the generalization of our findings to other cases. Section 6 (Conclusion) summarizes the key findings of the study and highlights limitations of the study along with suggestions for further research in the field.

# Theoretical Background
In this section, we review and discuss the theoretical background and previous literature related to our study. The research subject of our study has been approached from multiple different disciplines within both the humanities, natural sciences, and social sciences. In the following theoretical and empirical review, we include related research across different scientific approaches, while our own approach is mainly sociological.

The first part (2.1) describes how theories of the relationship between media and audiences have developed until today, and how they are relevant to our study. The second part (2.2) focuses on the consequences of social media and digital technologies and relates this digital evolution to the relationship between media and audiences. These two first parts serve as background information that places our study in a broader theoretical context. Subsequently, the third part (2.3) elaborates on the concept of audience engagement on social media and presents findings from previous studies. In the fourth part (2.4), we consider news content might be understood, and what characteristics of news content might influence audience engagement. In the final part (2.5), we discuss two empirical limitations within previous research, which shows the relevance of our current study.

## Media and Audiences
The relationship between media and audiences is an essential topic within media and communications research (Livingstone 2016:1). The concept of an audience has been central since the first formulations of media theories, and in the current digital age, the concept of audiences is more prevalent than ever (Livingstone 2016:1-2). In this section, we will describe the theoretical developments of the theories about media and audiences, and finally discuss how the idea of audiences has become central to contemporary debates in media studies.

### Mass Media and Passive Audiences
\noindent
The early formulations of theories about media are closely related to the emergence of mass media (Merrin 2014:95). The popularization of broadcasting through radio and television gave rise to the concept of a mass being stimulated with content by a single media producer (Merrin 2014:94). Lippmann (1922) was one of the first to describe the idea of a transmission model where media is sent from the producer to the audience. He emphasized how mass media was shaping public opinion by transmitting cognitive stereotypes to a mass of people (Lippmann 1922). Consequently, he was quite critical of mass media, as it carries the potential to manipulate rather than inform the public. This critical perspective has been continued even more famously in the works of Herman and Chomsky and their idea of manufacturing consent (Herman & Chomsky 1988), as well as in the works of Adorno and Horkheimer on enlightenment as mass deception (Adorno & Horkheimer 2002; Adorno 1991). In both of these approaches, the mass audience is deceived, as the media present them with simplified or biased portrayals of the world, promoting conformity rather than an informed and critical consciousness (Adorno 1991:104).

The concept of a transmission model of media has also been developed within a tradition that is less normative, adopting a more positivistic approach. Prominently, the two mathematicians Shannon and Weaver (1949) summarized the relationship between media and audiences in their sender-message-receiver model. This theoretical model has a clear causal relationship: a sender (e.g., the news media) transmits a message, which is received by the audience and produces a specific effect. Similar to Shannon and Weaver’s model, Lasswell (2007) also described a transmission model formulated as a chain of questions: who says what in which channel to whom with what effect (Wanta & Myslik 2019:59-60)? These theoretical models have had a large impact on media studies, and many studies have examined media effects of different kinds of content and audiences, but always with an emphasis on a one-way, causal direction from media producers to media consumers (Wanta & Myslik 2019:59-60).

### The Idea of Active Audiences
\noindent
Most contemporary research within media studies has moved away from the notion of audiences as passive masses without any agency. In the latter half of the 20th century, several different approaches contested the previous understanding of audiences. Different theoretical approaches have argued that audiences should be seen as being more diverse and having more agency. Three of the most important approaches that we will mention here are: 1) interpersonal relations by Katz & Lazarsfeld (1955), 2) the encoding-decoding model by Hall (1973), and 3) Livingstone’s (2005) concept of active audiences.

Katz and Lazarsfeld (1955) were among the first to challenge the mass-communication-idea of audiences as a passive receiver. They identify five different intervening variables affecting the relationship between media and audiences: exposure level, medium type, media content, audience predispositions, and interpersonal relations (Katz & Lazarsfeld 1955:20-25). Their work emphasizes the fifth variable, interpersonal relations, since it gives a new kind of agency among the audience members. Readers of a newspaper and viewers of a television shows are not just separate individuals, but they have interpersonal relationships through social networks and group norms, which mediates the influence of the media (Katz and Lazarsfeld 1955:43-44). Katz and Lazarsfeld also introduced the idea of a two-step flow of communication, where opinion leaders – people with especially significant interpersonal influence – distribute and mediate the flow of information from media to the audiences (Katz & Lazarsfeld 1955:32).

Another important contribution within media studies is the works of Hall (1973). He argues that there can never be one single stimuli-effect of a media message, because both the process of media production and media reception are embedded in ideologies (Hall 1973:265). When journalists and media workers produce content, it is encoded with different knowledge and assumptions, and subsequently the media content is decoded by the audience and reacted to according to their structural and cultural position.

In more contemporary research, the idea of active audiences has also been greatly emphasized by Livingstone (2006:9). She argues that audiences are increasingly diversified and situated within specific communities (Livingstone 2006:9-10). This perspective strongly emphasizes that media messages are fully interpreted within communities, and thus the approach has had a strong focus on ethnographic methods (Livingstone & Das 2013:1-2). In its most extreme formulation, there is no inherited meaning in media content, and media content will have completely different meanings for different communities (Philo et al. 2015:464).

Although the approaches of Katz and Lazarsfeld, Hall, and Livingstone are quite different in nature, they all share the common trait of imagining an audience that has some kind of agency – both towards media producers and each other. To some extent, the theoretical approaches to media and audiences can be placed on an axis of how much agency is attributed to the audience, and how unified or fragmented the audience is thought to be (Philo et al. 2015:464). While recognizing that these underlying dynamics are a part of the subject, our study will have a specific empirical focus on the relationship between the news content and audience engagement.

### Audiences 2.0
\noindent
The idea of an audience is both a central pillar in media theory and an essential part of everyday practices in newsrooms (Livingstone 2016:1-2). Nevertheless, many researchers have started to question the contemporary relevance of the concept of audiences. The critique is mainly a consequence of the concept being firmly connected to specific technological mediums: mass broadcasting media such as radio and television (Livingstone 2016:2). Recently, mass broadcasting media has been greatly challenged by digital media. Some argue that this technological evolution also has deep consequences for previous media theories - because they simply no longer fit the digital reality - to such a degree that we need to develop a media studies 2.0 (Merrin 2014).

While it seems exaggerated to discard decades of media studies, there is no doubt that the digital technological advancements have introduced new possibilities for the media ecology and not least for the relationship between media and audiences. Where traditional mass media is characterized by a one-to-many-relationship, digital media allows for a many-to-many-relationship (Livingstone 2016:2). On the internet, anyone can produce content that resembles journalistic content. This basic access to publishing challenges one of the most rudimentary principles of traditional journalism: that the journalist is the gatekeeper of producing, selecting, and exposing content to the audience, which passively consumes it (Deuze 2005:451). While there is a possibility for everyone to take up the job as a journalist, it remains questionable if everyone actually does involve themselves in producing content (Livingstone 2016:2). With technological breakthroughs, there is a tendency make overly enthusiastic proclamations of an entirely new era, but we have yet to see the distinction between news content producers and news content consumers completely vanish (Webster 2014:2; Livingstone 2016:2).

### Subconclusion: Media and Audiences
\noindent
This first part of the theoretical background outlines how the concept of audiences has been central to the interdisciplinary field of media studies. Today, most theories of audiences and media regard the audiences as active and diverse communities. The concept of audiences is even more emphasized and discussed with the invention of new, digital media, where audiences have even more agency. Our study continues in the line of research that understands audiences as active and participating by examining audience engagement behavior with news content on social media.

## Digital Media
In the previous section, we consider how the evolution of digital technologies and especially social media have had significant consequences for the relationship between media and their audiences. In this section, we describe these new digital technologies and news media in a broader sociological context. Lastly, we discuss previous research of the social media, Facebook, which is the specific focus of our study.

### Digital Trace Data
\noindent
When researchers refer to the digital age or the evolution of digital technologies in the 21^st^ century, the concrete technological inventions often concern the exponential increase in computational power and storage capacity (Salganik 2018:3-4). Every second, massive amounts of data are written to huge databases. Included in this stream of data is especially the monitoring of our digital and digitalized behavior – commonly named digital trace data. Digital trace data has several advantages. First of all, they are big i.e., they exist in large quantities (Salganik 2018:17). Furthermore, they are found data meaning that it is real, digital or digitized behavioral data, and therefore also non-reactive to the collection by the researcher (Salganik 2018:24). There are also some disadvantages of using digital trace data. These disadvantages are that digital trace data can be noisy, algorithmically confounded, or incomplete – meaning that relevant features such as background information of the users might not be available (Salganik 2018:35ff). The concept of big data has also been used to describe the accumulation and analysis of digital trace data. However, this term is often used in a variety of ways with no clear singular definition (Sætnan et al. 2018:6).

In the context of news and media, the evolution of digital trace data has also had a huge impact. In the literature, several studies have examined how audience analytics mainly based on these digital trace data have made its way into the newsrooms (Christin 2018; Ferrer-Conill & Tandoc 2018). The implementation of different analytics services and dashboards in the everyday of the journalists, editors, or even specific audience editors have been commonplace in many different media organizations (Ferrer-Conill & Tandoc 2018). Digital trace data and predictive modelling have become especially valuable for media organizations, since it is related to recommendation of advertisements, which is a major source of income for most modern media organizations (Martin 2019:21). Moreover, as the study by Christin (2018) of an American and a French newsroom have shown, digital trace data can be interpreted and used quite differently within media organizations (Christin 2018). The constant interpretation of performance metrics might cause a reactivity among journalists where they adapt their work as they become aware that they are being evaluated of their work (Christin 2018:1386; Espeland & Sauder 2007:1). In this way, the introduction of quantitative feedback based on the digital trace data can have both good and bad consequences, both in terms of making content more relevant to the audience, but it could also decrease the kind of news that are important but might not be among the most popular (Christin 2018:1409). While these studies focus on the effect on audience engagement on the media production, our study focuses on what precedes these evaluations: the measurement of how audiences engage with different news content.

### Social Media
\noindent
In addition to the technical evolution of the capacities of computers, a significant newcoming in the 21st century has been the invention and popularization of social media or social network sites. Social media has many different definitions, but a common trait is that it is a digital service that connects people (Aichner et al. 2021:220). Thus, social media is different from previous media, because it is social – it is many-to-many communication rather than one-to-many mass communication (Livingstone 2016:2). Previous research has identified different uses of social media such as socializing with friends or family, romance and flirting, job seeking or networking, and other kinds of interaction with strangers, organizations, companies, content and much more. The many uses of social media highlight how many different social phenomena have been digitalized or amplified by the invention of social media (Aichner et al. 2021:220).

A lot of research has been devoted to the study of social media, and a few areas of research stand out as particularly prevalent. This is especially evident for the phenomenon of social and political polarization, which has garnered significant attention in social media research. Polarization refers to a division between different groups of people or sets of opinions, and the research area is closely connected to related concepts such as echo chambers or filter bubbles that both refer to a social homophily of ideologies, i.e., that you only interact with other people, who have the same opinions as yourself (Bail 2021). A lot of studies have examined the prevalence of polarization of social media, as it has been hypothesized that political polarization is increased on social media. For example, Bail et al. (2018) find that Twitter users being exposed by statements of the opposite republican or liberal ideology become more extreme in their respective political ideology. In general, they find that users become less moderate in their opinion by being on social media and being exposed to different political statements (Bail et al. 2018; Bail 2021). However, the popularized concepts of echo chambers and filter bubbles have been quite contested, as they represent a too simplified description of the informational exchanges on social media (Bail 2021). For example, a study by Gentzkow and Shapiro (2011) found that the polarization of engagement with news media on social media is lower compared to offline media engagement such as newspaper subscriptions. Thus, polarization might be a problem, but the idea of social media amplifying polarization seems to be more complex than first anticipated. Similarly, Bakshy et al. (2015) found that individual selection when engaging with media content on social media leads to more polarization than the algorithmic sorting of the media content. Another interesting study is Schmidt et al. (2017), who finds that more active users tend to focus on a small number of news sources in their social media feed. Rather than constructing personalized filter bubbles of news content, current research suggests that the social media platform tend to homogenize the news content exposure, so that a small number of news media gain most of the exposure and engagement on social media (Nechushtai et al. 2023:18-19).

Another area of research within social media and news is the impact of misinformation, disinformation and fake news. The increase of misinformation and fake news is often related to polarization as studies have shown how strong political partisanship might drive believing in factual incorrect media content (Bail 2021:109). Previous research has examined how adding warning tags or correction statements might influence the effect of misinformation, but the studies are not in agreement (Muhammad & Mathew 2022:279). Some studies have found that warnings can reduce the spread of misinformation, while other have suggested a backfire-effect – a cognitive confirmation bias, where corrections of misinformation increase the belief in the incorrect information (Nyhan & Reifler 2010; Wood & Porter 2019).

There are many different social media with different specific functionalities, norms and likely also different types of people using them. Currently, social media such as Instagram, Snapchat and TikTok are most popular among younger people, while a social media like LinkedIn is more popular among those with higher levels of education (Auxier & Anderson 2021). The popularity of social media can change quickly, and previous popular sites such as Myspace or Google+ have completely disappeared from the landscape of social media. However, in most countries Facebook remains the most popular and widely used social media (Auxier & Anderson 2021; DR Medieforskning 2022:13; Ortiz-Ospina 2019).

### Facebook
\noindent
Across the world, Facebook remains the most popular social media with close to 3 billion daily active users (Meta 2022). In Denmark, 65% of the population use Facebook on a daily basis (DR Medieforskning 2022:13). According to Facebook itself, the mission of the company is *“Giving people the power to build community and bring the world closer together”* (Meta 2023). To a certain extent, the company’s slogan reflects that Facebook have always been a site for managing your social network, and previous research have also shown that the primary uses for joining Facebook are social interaction and social identity-building (Caers et al. 2013:984-985). However, Facebook and especially the company behind Facebook, Meta, have over time developed into an overarching platform by co-integrating with Meta’s many other social media and digital services such as Messenger, Instagram, WhatsApp (Helmond et al. 2019). Similar with most other social media, the financial model for Facebook is based on using digital trace data of the users to sell advertisements to over 90 million businesses around the world (Helmond et al. 2019).

Engagement with media and news content is also a vital part of using Facebook. In a Danish context, Facebook is the largest source of news content exposure compared to other social media (DR Medieforskning 2018:48). Especially for young people between 15 and 24 years old, Facebook serves as the primary source for news content (DR Medieforskning 2018:48). However, the relationship between Facebook and media organizations has been complicated. The most significant dispute between Facebook and news media was when the Australian government in 2019 tried to force Facebook to pay media organizations for showing their news content, and Facebook reacted by removing all news content for Australian users (Leaver 2021). While an agreement was eventually reached and Australian news content was reinstated on the Facebook platform, this incident highlighted the dual role of Facebook as both a distribution platform for news content exposure and a potential risk to the business models of media organizations. News organizations often share their content on Facebook, but the underlying goal is commonly to use social media engagement to direct potential audiences away from Facebook and onto their respective websites (Humprecht et al. 2020:1).

### Subconclusion: Digital Media
\noindent
This second part of the theoretical background describes how digital technologies and the invention of social media has had a significant impact on news media and their audiences. Extensive research has been dedicated to studying social media, where a particular focus has been on areas such as polarization and misinformation. Lastly, the relationship between audiences and news media is especially interesting on Facebook, as this social media is the largest and for many serves as a central source of engaging with news content.

## The Many Faces of Audience Engagement
In previous research, the idea of audience engagement has been defined and measured in many different ways. When studying audience engagement on social media, measurements of audience engagement have been everything from reading the news content, liking the post, sharing the article, commenting on the post, and not least the content and tone of the conversations around the article. In the following section, we will discuss how audience engagement can be defined, and how we may think of it as a multifaceted phenomenon in which different kinds of behaviors relate to different kinds of motivations for engaging with the news content.

### What is Engagement?
\noindent
The concept of audience engagement is growing in popularity within both academia and media organizations, and it has become almost a media industry buzzword (Nelson 2021:2350-2351; Lawrence et al. 2018). In some newsrooms, there are job descriptions focusing on audience engagement, and there is a growing industry of audience engagement services and tools offered to journalists and media organizations (Nelson 2021:2351). However, there is often a lack of concrete theoretical definitions of what engagement actually entails (Steensen et al. 2020:1664).

One very broad definition of engagement is that it *“refers to the cognitive, emotional, or affective experiences that users have with media content or brands“* (Broersma 2019:1). The main component of this definition is that engagement is different from simply exposure – it *“denotes an active and intentional orientation toward what users read, view, or hear.”* (Broersma 2019:1). However, this is still a quite open and perhaps unprecise definition of engagement (Steensen et al. 2020:1664). Some research additionally distinguishes between felt and behavioral engagement, where felt engagement refers to affective and cognitive reactions, and behavioral engagement means practices that can be observed (Steensen et al. 2020:1665). Thus, felt engagement could be having angry or happy feelings about content or changing attitudes as a result of content, while behavioral engagement could be reading a news article, liking content on social media, sharing news content, or writing comments.

In both academia and media organizations, there have been an emphasis on behavioral engagement, because it is easier to measure and quantify, and behavioral engagement might also be used as a proxy for felt engagement (Steensen et al. 2020:1665). However, even if only behavioral engagement is considered, we argue that there are substantial differences in what the different audience engagement behaviors represent. In their review on definitions of audience engagement, Gajardo and Meijer (2022) argue that previous research has predominantly concentrated on attention and time spent on news content, while less consideration has been given to how the time is spent or the quality of the engagement. They highlight audience engagement as a complex and multifaceted phenomenon composed of different dimensions of behaviors and emotional orientations (Gajardo & Meijer 2022; Steensen et al. 2020). In the following, we will further elaborate how different audience engagement behaviors might relate to various underlying motivations.

### Motivations for Engagement
\noindent
To understand different audience engagement behaviors, we argue that it is important to consider what motivation lies behind. Multiple studies have found that different behaviors such as liking, sharing, or commenting are related to different user motivations (Almoqbel et al. 2019; Kim & Yang 2017). In fact, there is a theoretical approach within media studies that emphasizes how different kinds of media uses are related to different kinds of gratifications, i.e., different sociological or psychological motivations (Ruggiero 2000). Here, the question of what constitutes engagement is not only the study of the kind or extent of engagement, but also what precedes or motivates the actual action of engagement.

The theory of uses and gratifications has been described in many ways, but originates from the work of Katz and Lazarsfeld (See Section 2.1.2). One particular sociological formulation is the theory by Levy and Windahl (1984), who stresses that gratifications should not be understood as the audience being superrational and selective (Ruggiero 2000:8; Levy & Windahl 1984:52ff). Levy and Windahl developed a taxonomy in which audience engagement can be placed within both the audience motivation and the phase of exposure: pre-, under-, or post-exposure of the news content. In this way, one kind of audience engagement might be related to a selective orientation, which happens pre-exposure of the news content. Other kinds of audience engagement such as discussing the news content is oriented towards socializing, and it happens after the exposure of the news content. In this way, they provide a theoretical example of how different kinds of audience engagement might be explained and related to different kinds of motivations.

The uses and gratifications approach might be particularly relevant in the case of audience engagement on social media, as this new form of media makes possible an increased agency of the audiences (Ruggiero 2000:20) (See also section 2.1.3). In line with this, a study by Springer et al. (2015) applies the approach to audience engagement on social media - specifically the behaviors of commenting and reading comments. They identify four dimensions of audience engagement, which relate to different underlying motivations. These dimensions include cognition, entertainment, social-integration, and personal identity, and they divide user behavior into commenting (e.g., commenting) and lurking (e.g., reading) behavior (Springer et al. 2015:800ff). Similarly, a review by Kümpel et al. (2015) identifies three broad motivations for online news sharing behavior: self-serving motives like promoting reputation or gaining social status, socializing and discussing ideas, and altruistic motives of information sharing or seeking (Kümpel et al. 2015:6).
